% Attributions for Poverty in Post-Socialist Countries
% Markus Kainu & Mikko Niemelä
% Nov 2013

`generated at:`
```{r, echo=FALSE}
Sys.time()
```
---

<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link> <link href='http://fonts.googleapis.com/css?family=Oswald' rel='stylesheet' type='text/css'>

```{r, eval=FALSE, echo=FALSE}
setwd("~/workspace/lits/attrib/attrib_year2013")
library(kaRski)
knitpandoc("index")

cd ~/workspace/lits/attrib/attrib_year2013/
lftp -u np16003 -e "mput *.html *.Rmd *.pdf *.epub *.docx;quit" ftp://np16003@www.muuankarski.org/domains/muuankarski.org/public_html/research/attributions/

cd ~/workspace/lits/attrib/attrib_year2013/figure
lftp -u np16003 -e "mirror -R;quit" ftp://np16003@www.muuankarski.org/domains/muuankarski.org/public_html/research/attributions/figure
```

```{roptsattrib, echo=FALSE}
opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=12)
```


`log of changes`

date      comment
--------  -------------------
20131103  Regression modelling in process. Final models will be published on Tuesday afternoon



# Introduction

This site gathers together all the data analysis behind the actual paper. Data analysis is broken down in three parts and each part is presented in .html format for human eyes and in text source including the R-code required for reproducing the analysis. As both the [micro data](http://www.ebrd.com/pages/research/publications/special/transitionII.shtml) and [macro data](http://www.qog.pol.gu.se/data/) are available as open data, anyone can rerun these analysis.

**Analytical code is licensed under open [FreeBSD](http://en.wikipedia.org/wiki/BSD_licenses#2-clause_license_.28.22Simplified_BSD_License.22_or_.22FreeBSD_License.22.29)-licence.**


# Data Analysis

## Cleaning the data & variable summaries

-------------------
html-ouput                                          source code
------------                                        --------------
[loading, cleaning and summarizing](loadClean.html) [source](loadClean.Rmd)
-------------------

## Descriptive analysis

-------------------
html-ouput                                        source code
------------                                      --------------
[Descriptive analysis](descriptive_analysis.html) [source](descriptive_analysis.Rmd)

[Spatial visualization](mapPlots.html)            [source]((mapPlots.Rmd))


[Graphics in the paper](finalPlots.html)          [source](finalPlots.Rmd)
-------------------


## Modelling the data

-------------------
html-ouput                                                source code
------------                                              --------------
[Regression analysis](regressionModelling.html)           [source](regressionModelling.Rmd)

[Final tables](finalTables.html)                          [source](finalTables.Rmd)
-------------------

- [Final tables as A4 .pdf](finalTables.pdf)

# Paper

- [Article draft in .html](article2013.html)
- [Article draft in .pdf](article2013.pdf)
- [Article draft in .epub](article2013.epub)
- [Article draft in .docx](article2013.docx)



# Presentation slides

- HSE Moscow - November 12, 2013
- Jatkokoulutusseminaari Turku November 7, 2013
- [ESA Torino - August 29, 2013](torino2013/TorinoSlides2013_final.html)


