% Attributions for Poverty in Post-Socialist Countries
% Markus Kainu & Mikko Niemelä
% Nov 2013

`generated at:`
```{r, echo=FALSE}
Sys.time()
```
---

<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link> <link href='http://fonts.googleapis.com/css?family=Oswald' rel='stylesheet' type='text/css'>

```{r, eval=FALSE, echo=FALSE}
setwd("~/workspace/lits/attrib/attrib_year2013")
library(kaRski)
knitpandoc("index")

cd ~/workspace/lits/attrib/attrib_year2013/figures
lftp -u np16003 -e "mirror -R;quit" ftp://np16003@www.muuankarski.org/domains/muuankarski.org/public_html/research/attributions/
```

```{roptsattrib, echo=FALSE}
opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=12)
```


# Introduction

This site gathers together all the data analysis behind the actual paper. Data analysis is broken down in three parts and each part is presented in .html format for human eyes and in text source including the R-code required for reproducing the analysis. As both the [micro data](http://www.ebrd.com/pages/research/publications/special/transitionII.shtml) and [macro data](http://www.qog.pol.gu.se/data/) are available as open data, anyone can rerun these analysis.

Analytical code is licensed under open [FreeBSD](http://en.wikipedia.org/wiki/BSD_licenses#2-clause_license_.28.22Simplified_BSD_License.22_or_.22FreeBSD_License.22.29)-licence.


# Data Analysis

## Cleaning the data & missing data

html-ouput                                                    source code
-------------------------------------------------------       -----------------------------
[loading and cleaning](loadClean.html)                        [source](loadClean.Rmd)
[missing data & diagnostics](code/descriptive_analysis.html)  [source](code/descriptive_analysis.Rmd)

## Descriptive analysis

html-ouput                                                source code
-------------------------------------------------------   -----------------------------
[Descriptive analysis](descriptive_analysis.html)         [source](descriptive_analysis.Rmd)
[Graphics in the paper]()                                 [source](descriptive_analysis.Rmd)

## Modelling the data

html-ouput                                                source code
-------------------------------------------------------   -----------------------------
[Output tables in html](code/multilevel_models.html)      [source](code/multilevel_models.Rmd)
[Output tables in pdf](code/multilevel_models.pdf)        [source](code/multilevel_models.Rmd)

# Paper

- [Article draft in .html](article2013.html)
- [Article draft in .pdf](article2013.pdf)
- [Article draft in .epub](article2013.epub)



# Presentation slides

- HSE Moscow - November 12, 2013
- Jatkokoulutusseminaari Turku November 7, 2013
- [ESA Torino - August 29, 2013](torino2013/TorinoSlides2013_final.html)


